---
title: "Data manipulation - solutions"
author: "[Gbadamassi G.O. Dossa](dossa@xtbg.org.cn)"
date: "Updated on `r Sys.Date()` (created on 2021-09-13)"
output:
  xaringan::moon_reader:
    css: [default, "slides-theme.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: ''
      titleSlideClass: [center, middle]
---

```{r setup, include=FALSE, echo=FALSE}
#options(htmltools.dir.version = FALSE)
#knitr::opts_chunk$set(comment = "", warning = FALSE, message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(palmerpenguins)
library(tidyverse)
```


class: center
## Acknowledgements
The content of this module are based on materials from:

.pull-right[
]
.pull-right[
[olivier gimenez's materials](https://oliviergimenez.github.io/)
]

---

## Question 1a 

.tiny-font[
```{r}
#read libraries
library(palmerpenguins)
library(tidyverse)
penguins # display data
glimpse(penguins)
summary(penguins)
```
]

---

## Question 1a 
a. Display the data `penguins`.  
.tiny-font[
```{r}
penguins %>% glimpse() # display data
```
]

---

## Question 1b
b. Make sure you understand the columns we have in this dataset.
c. Filter out penguins for which sex is missing. 
d. Select variables species, island, bill_length_mm and body_mass_g. 
e. Store the new dataset in a dat object. 

```{r}
dat <- penguins %>%
# filter out missing sex
  filter(!is.na(sex)) %>% 
# select variables
  select(species, island, bill_length_mm, body_mass_g) 

glimpse(dat)
```


```{r}
penguins %>%
# filter out missing sex
  filter(!is.na(sex)) %>% 
# select variables
  select(species, island, bill_length_mm, body_mass_g)-> dat2 


glimpse(dat2)
```
---

## Question 2a
a. How many penguins do we have in the dataset? 


.tiny-font[
```{r}
dat # nb of penguins
```
]

---

## Questions 2b and 2d
b. How many species? 
d. Count the number of penguins per species.

```{r}
# nb of species, and penguins per species
glimpse(dat)
levels(dat$species)
dat %>% count(species) 
summary(dat)
```

---

## Questions 2b and 2d
b. How many species? 
d. Count the number of penguins per species.

```{r}
dat %>% count(species, sort = TRUE) # idem, arranged by n
```


---

## Question 2c
c. How many islands? 

```{r}
dat %>% count(island) # nb of island, and penguins per island
```


---

## Question 2e
e. Count the number of penguins per species and per island.

```{r}
# penguins per species and island
glimpse(dat)
dat %>% count(species, island) 
```



---

## Question 3a: mean body mass
a. Calculate the overall mean body mass.

.tiny-font[
```{r}
# option 1
mean(penguins$body_mass_g) # Gives NA because the original data has missing values
mean(dat$body_mass_g)# here no more NA becuse we filtered out NA
a<-dat %>%
  mutate(mean_bm = mean(body_mass_g))
glimpse(dat)
glimpse(a)
a
```
]

---

## Question 3a: mean body mass


```{r}
# option 2
dat %>%
  summarise(mean_bm = mean(body_mass_g))
```


---

## Question 3b: mean body mass per species
b. Calculate the mean body mass for each species.

.tiny-font[
```{r}
# option 1
dat %>%
  group_by(species) %>% 
  mutate(mean_bm = mean(body_mass_g))
```
]



---

## Question 3b: mean body mass per species
b. Calculate the mean body mass for each species.
```{r}
# option 2
dat %>% 
  group_by(species) %>% 
  summarize(mean_bm = mean(body_mass_g))
```


---

## Question 3c: mean traits
c. Calculate the mean of both traits bill length and body mass measured for each species.

```{r}
# all at once, through column selection
dat %>% 
  group_by(species) %>% 
  summarize(across(bill_length_mm:body_mass_g, mean))
```



---

## Question 3c: mean traits

```{r}
# all at once, through column format selection
glimpse(dat)
dat3<-penguins%>%
  filter(!is.na(sex))%>%
  select(species, island, sex, bill_length_mm, body_mass_g)
glimpse(dat3)
dat %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean))
dat3%>%
  group_by(species)%>%
  summarize(across(sex:body_mass_g, mean))

dat3%>%
  group_by(species)%>%
  summarize(across(where(is.numeric), mean))
```
